### 2019-06-15 paper/paper.tex
added a few citations to the paper

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{chemmacros}
\usepackage{natbib}
\usepackage{lineno}
\usepackage[margin=1in]{geometry}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Solution of Differential Algebraic Equations with Local Constraints on the Latent Space of an Autoencoder},
            pdfauthor={Alejandro Francisco Queiruga},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
%\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% Patch equations to properly generate line numbers for review
\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
  \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
  \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
  \renewenvironment{#1}%
     {\linenomath\csname old#1\endcsname}%
     {\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
  \patchAmsMathEnvironmentForLineno{#1}%
  \patchAmsMathEnvironmentForLineno{#1*}}%
\AtBeginDocument{%
\patchBothAmsMathEnvironmentsForLineno{equation}%
\patchBothAmsMathEnvironmentsForLineno{align}%
\patchBothAmsMathEnvironmentsForLineno{flalign}%
\patchBothAmsMathEnvironmentsForLineno{alignat}%
\patchBothAmsMathEnvironmentsForLineno{gather}%
\patchBothAmsMathEnvironmentsForLineno{multline}%
}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\title{Solution of Differential Algebraic Equations with Manifold Constraints
on the Latent Space of an Autoencoder}
\author{Alejandro Francisco Queiruga}
\date{\today}

\begin{document}
\maketitle
\begin{abstract}
  Multiphase simulation is hard.

Our
proposed method is motivated by the problems of multiphase reaction and
transport, where the underlying constraints are the ill-behaved material
properties which are difficult to describe.

Our proposed approach is to build upon modern machine learning
techniques to aid in equation discovery. An autoencoder is trained on
the original material dataset to learn a lower-dimensional
representation of the underlying manifold. The model is designed to be
directly incorporated into a macroscale simulation of physical balance
laws, where now the traditional partial differential equations are being
solved on the learned latent space. This provides a more robust
numerical algorithm, while putting the onus of programming part of the
simulation itself onto the computer: tens of thousands of lines of human
written code consisting of empirical fits and branching logic are
replaced by code generated by the learned model. The methodology will
first be demonstrated on datasets based on thermodynamic intensive
variables, which are already challenging to analyze, before being
extended to richer datasets such as experimental images or molecular
dynamics. We stress that the proposed approach is only utilizing machine
learning on material data: the simulation is indeed solving physical
balance laws. Through deep learning approaches combined with
domain-specific goals, we can learn feature sets that are both
interpretable and directly apply to analyzing larger systems.

We train an autoencoder to determine the lower-dimensional
representation of the EOS surface using modern deep learning techniques
without using any phase labels. The decoder end of the model is then
plugged back into the balance laws, eliminating the constraint equation.
\end{abstract}

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\hypertarget{header-n3216}{%
\section{Introduction}\label{header-n3216}}

Representing complex physical systems with multiple speicies, phases,
chemical reactions, and heterogenous scale effects poses an
intractability problem to modeling. How do we analyze complex
experimental data and \emph{ab initio} simulations and derive new
macroscale relations to describe the systems? How do we then implement
these relations in the rest of our analyses and production-scale
computer simulations?

Accurately simulating multispecies reaction and transport with
undergoing phase transitions is a critical aspect to a variety of
problems.
This work is motivated by advancing simulation techniques for subsurface energy systems such as shale
and oceanic natural gas reservoirs in shales or engineered geothermal
systems.

The system of equations we are looking at are, for only two components,
\begin{align}
  \partial_t \rho_{H_2O} =& \nabla \cdot F_{H_2O} + s \\
  \partial_t \rho_{CO_2} =& \nabla \cdot F_{CO_2} + s \\
  \partial_t e =& \nabla \cdot F_{heat} + s
\end{align}
More species add more balance laws (rows of equations). The applications also involve continuum
mechanics and chemical reactions, which are not considered. The interplay of deformations,
reactions, and phase transitions makes the equations extremely difficult
to solve. \cite{queiruga2018millstone}
(In this paper, we use $\partial_t$, $\mathrm{d}/\mathrm{d}t$,
and $\dot{()}$ to denote differentiation with respect to time
interchangeably depending on context.
The complexity is hidden inside the relationships between temperature
$T$, and pressure,$p$, density, $\rho$, and internal energy $e$ (or
enthalpy, $h$, s.t. $e=\rho h - p$), 
for the materials in question.

% {\bf RESERVOIR PHYSICS LIT REVIEW HERE; Other applications;
% plasticity, }
One of the simulators with the widest-encompassing physical modules is
the TOUGH family of solvers that have been developed at Lawrence
Berkeley National Lab.
The methodology is based on a primary variable switching method in 
The simulator of THAT GUY WITH DUAL VARIABLES I MET AT AGU
parameterizes a simplified hydrate system using density and enthalpy
as the primary variables.
Higher order representations of flow are also a problem to tackle
anisotropy and increase the computational efficiency of the spatial
discretization. CITE ME AND MICKEY AND STEPHAN
Multiphase flow was done with finite elements in \cite{YANG2014417}.
THAT MORIDIS FEM PAPER
However, the complexity of the state representation restricts
development to a 2-phase

A single material with multiple phase transitions
is a challenge in of itself, and the descriptions become combinatorially
more complex as species and phases are added. This results in
combinatorially more human labor and combinatorially more code which
tends to involve complex branching logic. The descriptions then pose
numerical challenges when incorporated into macroscale simulations due
to obstacles such as discrete phase state machines and primary variable
switching.
This work only considers the "single beaker" problem; the flow
and transport aspects have similar challenges. (The next stage of this
work will "connect the beakers".)

We redescribed our problem as a differential algebraic equation
with a constraint: the balances of mass and energy are constrained by
the EOS. We can cast the above equation into a basic form for a
differential algebraic equation,
\begin{align}
  z( u, \partial_t u) &= 0 \\
  c(u) & = 0
\end{align}
where, for the equation above, the variable is $u=\{T,p,\rho,h\}$. We will introduce a new way of thinking about constraints as sets of
solutions or observations of possible $u$:
\begin{equation}
c(u)=0 \longleftrightarrow \left\{ u | c(u)=0 \;\text{or}\; u\; \text{observed} \right\}
\end{equation}
This maps directly to the original experimental procedure for
determining a constraint equation.

Because the EOS was originally derived from data, a data-driven
approach lends itself in which the EOS is a set of points in
temperature-pressure-density-enthalpy space. (In the multicomponent
cases there are more dimensions.)
We will use this set of possible values viewpoint to train an
autoencoder on the set of $u$ for a latent space parameterized by $q$:
\begin{equation}
u = D(q), \; q=E(u) \;\forall u \in U
\end{equation}
which satisifies
\begin{equation}
  D(q)\in U \forall q.
\end{equation}
We can then write the original equation by
\begin{equation}
z(q,\partial_t q)=0
\end{equation}
The actual meaning of the parameterization is not important, so long as
it can be effectively solved and then \emph{decoded} back to the
physically meaningful quantities.

% {\bf ML LIT REVIEW HERE} % TODO
\cite{wu_physics-informed_2018}
\cite{xie_tempogan:_2018}

 A program synthesis approach was used to search for and evaluate possible methods of Peridynamics in Queiruga, 2017 \cite{queiruga2017numerical}. (An open source implementation of the study can be found Periflakes\cite{queiruga2018periflakes}.)

Part of the new modern software engineering framework involves replacing
human-written code with computer-written code. We are applying this
mindset from industry problems in the Artificial
Intelligence space into scientific computing.

This work proposes a new methodology that steps away from the
traditional approaches. 
In this approach, we directly \emph{look for solvable representations}
instead of representations that carry human intuition. We value
simulation runtime and robustness. Further, we must remember the value
of the human effort involved: beyond improving computational aspects,
we are fully automating part of the human involvement in programming
the simulator.

We demonstrate this approach to solving constrained differential
equations on the pendulum, before demonstrating two EOSes of water: one
with just liquid-gas phases, and one with liquid-gas-ice-supercritical
regimes. The notion of phases does not appear in this multiphase
simulation. We show a model architecture that is able to learn phase
labels in the unsupervised autoencoder training process, such
human-scientist interpretable results are obtained from the machine
learning process. We compare this and other model architectures,
including common deep neural networks, based on autoencoder error and
simulation success metrics. We stress that the presented approach is
only utilizing machine learning on material relations derived from
empirical data: the simulation is indeed solving physical balance laws.

We build a test dataset off of the IAPWS solutions. The proposed
methodology applies to experimental data directly, or preexisting
compendiums of empirical fits and theoretical relations.
Note that, even if a relation can be theoretical derived, that does
not mean that the analytical relationship yields the best program
implementation: it is often the case that numerial approximations are
superior to closed-form mathematical solutions in computational
efficiency, accuracy, and numerical stability.

In section 2, the balance of mass and energy will be described for a
single multiphase material. We will discuss the current state of the
art algorithm for deriving relationships for multiphase materials, and
then the primary-variable switching algorithm that is used in
state-of-the art solvers. The equation will then be recast as a
constraint problem in th above form.
In section 3, we will demonstrate applying the autoencoder learning
methodology to solve the pendulum,
which can be considered the archetypical constraint problem.
In section 4, the autoencoder methodology wil lbe applied to the
equation of state of water. In section 5, the new simulation process
will be described, wherein the learned decoders are placed into the
balance laws, forming newly discovered equations on learned
representations. The software implementation of the training and
simulation programs are detailed in section 6. The new simulation
framework on sets of learned representations is evaluated on
multiple datasets of varying extents of water with a
hyperparameter search on a suite of test problems to prouce the discussion of section 7, as
well as a systematic automated testing procedure of a
machine-dominated programming methodology.
Active research and software development directions are
concluded in the final section.

The notebooks and software implementations are released open source to
accompany this work. They can be found at the github repository
located at: \href{https://github.com/afqueiruga/latentsim}.
Certain components are factored into submodule repositories which are
each released under the LGPL at seperate repositories referenced in
the implementation details.

\section{Multiphase Equations of State}

% \begin{figure}
%    CSTR
%   \caption{ \label{fig:cstr} Illustration of the Continuous stirred-tank
%   reactor. The material in the reactor is exposed to a reservoir with pressure
%   $p_\inf$ and temperature $T_\inf$ via a conduit with permeability
%   and thermal conductivities $k_p$ and $k_T$, direct forced mass
%   injection/withdrawl $r$, or heating and cooling $s$.}
% \end{figure}

For this preliminary work, we consider only a single-beaker problem
without flow, a continuous stirred-tank reactor (CSTR), as illustrated
in Figure \ref{fig:cstr}. Let \(\rho\) be the material density, \(T\) be the
temperature, \(p\) be the material pressure, \(h\) be the material
enthalpy, and \(u\) be the internal energy. There are two equations for
balance of mass and energy.
\begin{align}
\partial_t \rho & = \mathbf{k}_p(p_\infty - p) + r\\
\partial_t \rho u & = \mathbf{k}_T(T_\infty-T) + s 
\end{align}
The beaker is exposed to a reservoir with \(p_\infty\) and \(T_\infty\)
via a mixed boundary condition straw with transport and conduction
coefficients \(\mathbf{k}_p\), \(\mathbf{k}_T\). Mass and heating
sources and sinks are the terms \(r\) and \(s\).\footnote{We express the
  material by enthalpy instead of internal energy; they are related by
  the relation \(\rho u = \rho h - p\). Future work may work directly
  with internal energy.}
We are neglecting upstreaming in the connection between the reservoir
and the container.

We also consider the phase equilibria: beyond the scale of resolving
individual bubbles, it is possible
for a control volume to be a mixture of two phase coexisting.
(For a very fine scale simulation, e.g. at the pore scale, bubbles
would be explicitly modeled.)
We introduce a saturation $S_{phase}\in[0,1]$ for each phase;.
The material can exist on this section of the surface as a mixture of
the two phases. Thus, its density is calculated by a mixing the
properties of all phases at that point in the phase diagram:
\begin{equation}
  \rho = S_{gas} \rho_{gas} + S_{liq} \rho_{liq} + S_{ice}
    \rho_{ice} + S_{crit} \rho_{crit}
  \end{equation}
Each phase
and equilibrium surface constrains the possible saturations. For example, on the boiling equilibrium surface there are two phases with some mixture
saturations \(S_{gas}=1-S_{liq}\) and $S_{ice}=S_{crit}=0$. 
For water, three phases can exist in equilibrium. Given more species, it is possible for an arbitrary number of phases
to be in equilibrium; there is a quadruple point in methane-water
hydrates.

But, what is a phase boundary? The traditional explanation is shown in
Figure \ref{fig:phaseboundaries}. The use of intensive properties breaks down:\footnote{https://en.wikipedia.org/wiki/Phase\_transition}

\begin{quote}
Phase transitions occur when the thermodynamic free energy of a system
is non-analytic for some choice of thermodynamic variables (cf. phases).
\end{quote}

We can also assemble the surface of possible $T,p,\rho$ coordinates,
shown in Figure \ref{fig:pTrho}.
\begin{figure}
\centering
\includegraphics[width=4in]{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/figures/phase_diagram.png}
\caption{\label{fig:phaseboundaries}Phase boundaries for water on the $T-p$ diagram. The standard
interpretation is that the properties of the material change
drastically when crossing this boundary. Note how it is possible to form a path from a
definitively gas state to a definitively liquid state without crossing
a phase boundary. The supercritical regime has no clear boundaries,
but requires a seperate curve fit in practice due to numerical challenges.}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=4in]{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/figures/water_eos.png}
\caption{\label{fig:pTrho} The $\log{p}$, $T$, $\rho$ equation of
  state surface for water in units of $\log{\text{Pa}}$, K, and \si{}{kg/m$^2$} This plot spans four phases, three
  equilibria, and the triple point, which are each given their own color.
  There is a fourth dimension to the surface of enthalpy $h$ that exhibits
  similar discontinuities; the $\log{p}$, $T$, $h$ surface is plotted
  in the Appendix. (Every point on this plot is a point in the
  generated dataset used for the optimization)}
\end{figure}

{\bf Remark:}There are even more known phases of crystal ice which for water
For water, the supercritical phase and exotic ices only occur in
laboratory conditions or naturally near the
crust \cite{}. For our case, we only handle the Ice I and cutoff the
regime at \si{10^8}{Pa}. For other materials these exotic phases can be
applicable. For example, in geoengineered systems using $CO_2$
such as geothermal plants are carbon sequestration facilities, the
working fluid is usually in the supercritical regime at depth. cite
curt. cite hydrates.
When binary mixtures are involved, the number of possible phases and
equilibria increases again.

{\bf Remark:} This interpretation breaks down with the existenceof super-saturated phases in non-stable regions: e.g., it is possible to
heat water beyond it's boiling point in a microwave. We are concerned
with environments where this would not happen (the very rough and
heterogenous) pores of rocks, so we have not thought of how to
integrate this problem.

\hypertarget{header-n3253}{%
\subsection{State Machine logic}\label{header-n3253}}

The typical methodology is to use the empirical relations for density
and enthalpy \(h\) as a function of pressure and temperature and solve
the DAE for \(p\) and \(T\) implicitly:
\begin{align}
\text{Solve for}\, p(t)\, \text{and}\, T(t)\, \text{such that:}\\
\partial_t \rho(p,T) & = \nabla \cdot \mathbf{k}\nabla p + r\\
\partial_t \rho u(p,T) & = \nabla \cdot \mathbf{k'}\nabla T + s
\end{align}
The complicatation is that the functions \(\rho(p,T)\) and \(u(p,T)\)
are not well defined functions due to the presence of phase changes that
yield sharp discontinuities in \(p,T\), as shown in Figure 1 for
water.

\begin{tabular}{l|c|c}
  State & $X_1$ & $X_2$ \\
  \hline
  Ice & $p$ & $T$ \\
  Liquid & $p$ & $T$ \\
  Gas & $p$ & $T$ \\
  Supercritical & $\rho$ & $T$ \\
  Ice-Liquid & $S_{ice}$ & $T$ \\
  Ice-Gas & $S_{ice}$ & $T$ \\
  Gas-Liquid & $S_{gas}$ & $T$ \\
  Ice-Gas-Liquid & $S_{ice}$ & $S_{gas}$ 
\end{tabular}

  This requires a statemachine. For the complete states of water, it is shown in Figure \ref{fig:statemachine}.
  \begin{figure}
    \includegraphics[width=4in]{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/slides/phase_4.png}
    \caption{\label{fig:statemachine}States and possible transitions for water.}
\end{figure}
Formulate balance equations as
\begin{equation}
R(X ; phase) = 0
\end{equation}
Solve the differentiable part:
\[\frac{\partial R}{\partial X}\Delta X^{k+1/2} = -R(X^k ; phase^k)\]
and iterate on the nondifferential part:
\[phase^{k+1},X^{k+1} = statemachine(X^{k+1/2} ; phase^k)\]
The phase switching logic has the form of a switch case:

\begin{Shaded}
\begin{Highlighting}[]
\ControlFlowTok{switch}\NormalTok{ phase_old:}
  \ControlFlowTok{case}\NormalTok{ gas:}
    \ControlFlowTok{if}\NormalTok{ p,T crossing boundary:}
\NormalTok{      phase_new = liquid_gas}
  \ControlFlowTok{case}\NormalTok{ liquid:}
    \ControlFlowTok{if}\NormalTok{ p,T crossing boundary:}
\NormalTok{      phase_new = liquid_gas}
  \ControlFlowTok{case}\NormalTok{ liquid_gas:}
    \ControlFlowTok{if}\NormalTok{ S_gas >= }\DecValTok{1}\NormalTok{:}
\NormalTok{      phase_new = gas}
    \ControlFlowTok{if}\NormalTok{ S_liquid >= }\DecValTok{1}\NormalTok{:}
\NormalTok{      phase_new = gas}
\end{Highlighting}
\end{Shaded}

\hypertarget{header-n3262}{%
\subsection{A constrained Differential Algebraic
Equation}\label{header-n3262}}

We argue that the relations that are used are not the \emph{ground
truth} in of themselves. Beyond ideal gases and simple linear fluids,
the functions \(\rho(p,T)\) or \(p(\rho,T)\) are simply complicated fits
that were derived in an ad-hoc piecewise fashion from experimental data.
This is not a bad thing---it is the fundamental nature of the problem.
When we look at the corpus of literature on a material as water, what we
actually have is a decision branch to multiple different complicated
fits for each material branch.

We instead replace our problem with two balance laws and one unknown
constraint:
\begin{align}
\text{Solve for}\, \rho(t), \, h(t), \, p(t),\, \text{and}\, T(t)\, \text{satisfying:}\\
\partial_t \rho &  = \mathbf{k}_p(p_\infty - p) + r\\
\partial_t (\rho h-p) & = \mathbf{k}_T(T_\infty-T)  + s + h\left(\mathbf{k}_p(p_\infty - p) + r\right) \\
\text{such that they lie on the material EOS,}\\
\rho,h,p,T & \in \{\text{EOS}\}
\end{align}
The \emph{ground truth} is the experimental data in the first place; the
branching curve fit is only one realization of representing the data. We
have demonstrated that it is not possible to represent any two of these
in terms of the other two; and even if so, it is not necessarily a good
and robust fit. What if we search for a new fit?

\hypertarget{header-n3267}{%
\section{The Pendulum}\label{header-n3267}}

An archetypical constraint problem is the pendulum. Posed as dynamics
problem with a Lagrange-multipler-esque centripetal force, it
has three unknowns wtih two second-order equations and a constraint,

\begin{align}
\text{Solve for}\, x(t), \, y(t), \, f(t) \, \text{satisfying:} \\
m \ddot{x} & = f x/\sqrt{x^2+y^2} \\
m \ddot{y} & = f y/\sqrt{x^2+y^2} - m g \\
x^2 + y^2 & = R^2
\end{align}
Note that we are considering the nonlinear pendulum.

There are many ways to solve this cannonical equation: Lagrange
multipliers (\(f(t)\) is the multiplier in the way we have written it
above), penalties, change of variables, etc. The change of variables
method exploits the Lagrangian mechanics formulation, where we state
the Lagrangian,
\begin{equation}
L = \frac{1}{2}m\left(\dot{x}^2 + \dot{y}^2\right) - m g y
\end{equation}
with the constraint that \(x\) and \(y\) lie on the path.
We can introduce a new variable \(\theta\) that parameterizes
\(x=R\cos\theta\) and \(y=R\sin\theta\) and rederive the equations of
motion with the familiar,
\begin{align}
\text{Solve for}\, \theta(t) \ \text{satisfying:} \\
\frac{\mathrm{d}}{\mathrm{d}t} \frac{\partial L}{\partial
  \dot{\theta}} =
  \frac{\partial L}{\partial \theta}
\end{align}
In this example, this lets us collapse the problem into a
single-component second-roder differential equation.

In our methodology, we do not have an explicit equation for the
constraint, but a dataset of $X$ of $(x,y)$ pairs that are on the manifold. For
our proof-of-concept test, we manufacture this data for the circular
pendulum to verify against the analytical solution:
\begin{equation}
X = \left\{ (R\cos\theta_i, R\sin\theta_i)\; | \;\theta_i \in [-5\pi/4,\pi/4] \right\}
\end{equation}
We also try a ``goofier'' looking curve with no known closed form
solution to exercise the methodology,

(For the equation of state problem, this is analogous to the dataset
of $p,T,\rho,h$ 4-tuples that make the plot in Figure \ref{fig:pTrho}.)

The manifold for this simple problem is smooth, so two nested
polynomials forms the autoencoder:
\[\left\{\begin{array}{c} x\\y\end{array}\right\}
\rightarrow W_{enc} x^n +b_{enc}\rightarrow q \rightarrow W_{dec} x^n +b_{dec}\rightarrow 
\left\{\begin{array}{c} x\\y\end{array}\right\}\]
where $x^n$ denotes a polynomial basis set, e.g. for two variables $x^n=\{x,y,x^2,xy,y^2,..y^n\}$.
(We explicitly write the bias \(+b\); it is equivalent to including
\(x^0=1\) in the polynomial set.)

We'll use the autoencoder to generate the parameterization of \(x(q)\)
and \(y(q)\). To do this, we optimize the loss function,
\begin{equation}
L(x) = \left\| x-D(E(x)) \right\|
\end{equation}

When we want to solve the dynamics, we can treat the autoencoder the
same as \(x(\theta)\), plugging in
\[L(x(q),v(q,\dot{q}))\]
Taking the partial derivatives with respect to the feature variable and
its rate, the components to the equation of motion are build the DAE we know from physics:
\[\frac{\mathrm{d}}{\mathrm{d}t}\frac{\partial L}{\partial \dot{q}} = \frac{\partial L}{\partial q}\].
By the chain rule, the velocities are related to the
feature space by
\begin{equation}
\dot{x} = \frac{\mathrm{d}x(q)}{\mathrm{d}t} = \frac{\partial x}{\partial q}\frac{\mathrm{d}q}{\mathrm{d}t}
\end{equation}
and similarly for the \(y\) direction.
Then plug it into a Runge Kutta (trapezoidal) to solve with Newton's
method:
\[\frac{\partial L}{\partial \dot{q}}_i - \Delta t \frac{\partial L}{\partial q}_i = \frac{\partial L}{\partial \dot{q}}_0 + \Delta t \frac{\partial L}{\partial q}_0\]

Soltutions are shown in Figure \ref{fig:pendulum}.
\begin{figure}
\centering
\includegraphics[width=5in]{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/figures/pendulum_q.png}
\caption{\label{fig:pendulum}Solution of the pendulum on an arbitrary latent space.}
\end{figure}

\hypertarget{header-n3294}{%
\section{Representation Learning for
Reparameterization}\label{header-n3294}}

The central idea of this work is to obtain two new variables solve for
those directly which have no physical meaning other than simply being a
parameterization of the equation of state constraint. We want to select a
\(q_1\) and \(q_2\) to require no auxilary phase index to define the system
and no additional logic in the code. The sharp kinks and possible
discontinuities can be a part of the functions
\(\rho(q_1, q_2), p(q_1, q_2), h(q_1, q_2)\) and \(T(q_1, q_2)\).

\hypertarget{header-n3299}{%
\subsection{Latent Space}\label{header-n3299}}

Instead of the painstaking work of parameterizing phase boundaries and
determining curve fits by hand, we rephrase the problem of representing
the equation of state by learning and autoencoder. We have an encoding
phase, \(E(\rho,p,e,T; a)\) and a decoding phase, \(D(u,v; b)\) that
forms an identity function with a compressed subspace:
\begin{equation}
\left\{ \begin{array}{c}
T\\ p\\ \rho\\ h
\end{array}\right\} \rightarrow  E \rightarrow 
\left\{ \begin{array}{c} q_1\\q_2 \end{array} \right\}\rightarrow D \rightarrow 
\left\{ \begin{array}{c}
T\\ p\\ \rho\\ h
\end{array}\right\}
\end{equation}
The autoencoder is solved for by optimizing its parameters using the
goal

\begin{equation}
\min_a \sum_x \left( x - D(E(x;a);a) \right)^2
\end{equation}

Note that we are not yet considering constraints \(c\) with partial
differential equation components in space. These types of material
equations of states are only enforced pointwise. The time components of
the equations can contain spatial derivatives; i.e. this method fits
easily inside a finite volume simulation with little change.

\hypertarget{header-n3305}{%
\subsection{Training}\label{header-n3305}}

We use a Euclidean norm loss function with a contractive penalty:

\[L(x)=\left\|x-D(E(x))\right\|^2_2+\lambda\left\|\frac{\partial E}{\partial x}\right\|_F^2\]

The gradient on the encoder is a \(4\times2\) matrix in our case, and
\(\|\|_F\) denotes the Frobenius norm. The contractive penalty
\(\lambda\) smooths out and evenly distributes the mapping to \(q\).

The variational autoencoder is a little more complicated to implement,
and we do not necessarily want to match a Gaussian interior. Because we
generated our datasets artificially (both in this manufactured setting
and in an actual laboratory setting), the dataset does not reflect draws
from a distribution, so the KL divergence against a gaussian is not
quite applicable.

We note that penalizing the decoder would smooth out the sharp kinks at
phase boundaries.

The dataset is split into a standard $2/3-1/6-1/6$ division into
training-testing-validation groups. We note that we can always
manufacture more datapoints. We train by alternating phases. The first
step is minibatched stochastic descent on the entire deep network, assembling the losses on minibatches of the data, $X_{mini}$,
\begin{equation}
  L(X_{mini}) = \sum_{x\in X_{mini}} L(x)
  \end{equation}
to perform the out-of-the-box ADAM optimizer on all of the
parameters in the networks. (The box is
TensorFlow.) After a few epochs of this,
the training algorithm alternates to a few steps step of linear regression on the last layer of the network,
denoted by \(W^{dec}\), which directly controls the magnitudes of the values.
(The bias $b^{dec}$ is included in $W^{dec}$.) The linear
system that needs to be solved is easily generated by assembling the
gradient and Hessian of the total loss function,
\begin{align}
 \mathbf{G}\left(X_{train}\right) &= \sum_{x\in X_{train}}\frac{\partial L}{\partial W^{dec}} \\
\mathbf{H}\left(X_{train}\right) &= \sum_{x\in X_{train}}\frac{\partial^2 L}{\partial (W^{dec})^2}
\end{align}
and solving the linear system for an update to the last parameters.

\[[\mathbf{H}] \Delta W^{dec} = -\{\mathbf{G}\}\]

To form the hessian and gradient for this step, we sum over all of the
training data. (The contractive penalty term drops out of the this step because it does
not depend on \(W^{dec}\).)

\begin{equation}
\bar{H}_{ij} = H_{ij} + \left(\delta_{ij}\,\text{if}\,H_{ik}=0\forall k\right)
  \end{equation}

For this step in which only the back-end regression is performmed, each set of coefficients for the curves for each variable is independent from those of the other variables: e.g., the coefficients to the sets of polynomials that parameterize $p(q)$ do not depend on those for $T(q)$. This means that this step can be performmed in blocks, performing the linear step for each component independently, at a great reduction of cost. For the full system, 3/4 of the entries are zero. We did not bother performing this optimization, but note that the savings are quadratic as the number of coefficients grow for more complex equations of state. (This is not true for training the entire deep network: it is only true for this linear regression step on the sub-network.)

In future work, we would want to continue appending new \(q\)s in a
systemic way to derive, e.g., an \(H_2O + NaCl\) equation of state with
3 degrees of freedom, bootstraping from the pre-trained \(H_2O\)
architecture with 2 degrees of freedom.

\hypertarget{header-n3312}{%
\subsection{Encoder Initializations}\label{header-n3312}}

The properties \(p,T\) are the usual prefered choice for primary
variables, but as discussed above, the equations of state are not
analytic for any choice. (Indeed, this idea first spawned when the
author was thinking about systematically rewriting equations in
\(\rho,h\).) Ideally, this system would be fully automated, but due to
the nonlinearilty of the autoencoder problem there are many poor local minima.
Multiple starts are possibly needed
to get a good solution when starting from purely randomized newtork
parameters. Global optimization is truly a necessity when trying to
optimize networks with so few parameters. We incorporate a little bit of
human intuition and preference by trying three options to start off the
autoencoder: 1) fully random, 2) \(p,T\) prefered, 3) \(\rho,h\)
prefered. This is implemented by added a bias to the initialization of
the encoder coefficients, e.g. for \(p,T\),
\begin{equation}
 W^{init}_{enc} = \left[\begin{array}{ccccc}
1 & 0 & 0 & 0 & ... \\
0 & 1 & 0 & 0 & ...
                        \end{array}\right]+[\sigma]
\end{equation}
where \([\sigma]\) is the random initialization and \(…\) denotes
padding the coefficients to the nonlinear terms by 0s. Recall that the
values for \(s\) have already been rescaled to be around 0 with a range
of 1. The random initialization part is decided by
\(\sigma \sim \mathcal{N}(0,0.1)\) .

\hypertarget{header-n3317}{%
\subsection{Unsupervised Phase Classification}\label{header-n3317}}

The architecture is illustrated in Figure \ref{fig:classifyingnetwork}.
It has the following form in index notation:
\begin{align}
  q_i &= W^{enc}_{ij} x^n_j + b^{enc}_i\\
  f_{ij} &= W^{dec}_{ijk} q^m_k + b^{dec}_{ij} \\
  h_i &= \mu\left(W^{bound}_{ik}q^m_k+b^{bound}_{i}\right) \\
  p_i &= \sigma\left(W^{phase}_{ij}h_j+b^{phase}_i\right) \\
  \bar{x}_i&= f_{ij} p_j
\end{align}
where $x$ is the original prediction, $x^n$ denotes a polynomial basis
set on $x$ (which increases the number of channels), $q$ is the latent
space, $f$ is a set of polynomial curves, $h$ is a hidden layer
denoting sides of boundaries, $p$ is a classification layer depending
on $h$, and $\bar{x}$ is the predicted feature. The activation $\mu$
is chosen as a standard neuron, e.g. $\tanh$ or $\text{sigmoid}$. The
activation $\sigma$ is a probability activation, which is chosen to be
a softmax,
\begin{equation}
\sigma_i(x;\beta) = \frac{e^{\beta x_i}}{\sum_j e^{\beta x_j}}
\end{equation}
or, in the limit, the hard argmax:
\begin{equation}
\sigma_i \rightarrow \text{onehot}\arg\max z as  \beta\rightarrow\infty
\end{equation}

\begin{figure}
\centering
\includegraphics{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/slides/classifier_network.png}
\caption{\label{fig:classifyingnetwork}The architecture of the classifier network designed for
  unsupervised labeling of phases.}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=3in]{/Users/afq/Documents/Dropbox/ML/primaryautoencoder/slides/classifier_good.jpg}
  \caption{\label{fig:classified}Result of the unsupervised phase classifier. Compare this plot to the original labeled dataset Figure \ref{fig:pTrho}.}
\end{figure}
Because the identity of the phases,XXXXXXXX
For comparison purposes, we group together liquid-gas-supercritical as one contiguous label.

\hypertarget{header-n3321}{%
\section{Simulation}\label{header-n3321}}

Examining the $T,p,\rho$ surface allows for human interpretability, but
the decoder actually yields four new functions with two unknowns:
$D(q)=\{T(q),p(q),\rho(q),h(q)\}$. These surfaces can be plotted as
independent 2D functions: the surface of \ref{fig:classified} is
unfolded into the new parameterization in Figure \ref{fig:foursurfaces}.
\begin{figure}
  \centering
  \includegraphics{../slides/water_four_surfaces_plot.png}
  \caption{\label{fig:foursurfaces}The four 2D surfaces that make up
    the new parameterization. The learned surface captures the sharp
    kinks and vertical fronts in Figure \ref{fig:classifyingresult},
    but is an analytic and smooth function of the two learned unknowns.}
  \end{figure}

With the trained model, we solve for \(q_1(t)\) and \(q_2(t)\) such that:
\begin{align}
\partial_t \rho(q_1,q_2) & = \mathbf{k} \left( p_\infty-p(q_1,q_2) \right)  + r \\
\partial_t \left(\rho h(q_1,q_2)-p\right) & =\mathbf{k}_T(T_\infty-T(q_1,q_2))  + s + h\left(\mathbf{k}_p(p_\infty - p) + r\right)
\end{align}
where the expressions for  \(\rho(q_1,q_2)\) etc. are the back of the
trained autoencoder,. The trained
autoencoder automatically removes the local constraint, satisfying
\begin{equation}
eos(D(q_1,q_2)) = 0 \, \forall \,q_1,q_2
\end{equation}
Thus, the system will always satisfy the constraints as $q$ is
evolved, even during intermediate iterative solution steps. Note again
that the machine learning is only applied to the
material representation: the balance laws are expressed directly.

To develop the rest of the discretization, we introduce the compact form:
\begin{align}
\frac{\mathrm{d}}{\mathrm{d}t} m(T,p,\rho,h) &= r(T,p,\rho,h)\\
eos(T,p,\rho,h) &= 0
\end{align}
where $m=\{\rho,\rho h -p\}$ and $r$ is the right hand side, with no good representation of \(eos\) everywhere.

The decoder part of the network is inserted into the balance law
\begin{equation}
\frac{\mathrm{d}}{\mathrm{d}t} m(D(q)) = r(D(q))
\end{equation}
We can parameterize this equation with any time stepping scheme, .e.g. a
Runge-Kutta scheme. For subsurface flow, backward Euler is popular due
to its simplicity and L-stability, which we follow since the topic of
this paper is not exploring new time stepping schemes. This yields the
discretizated algebraic equation for \(q_i\) and \(t+\Delta t\) given the
value \(q^0\) at \(t\),

\begin{equation}
0 = R(q^i) = m(q^{i}) - m(q^{0}) - \Delta t \, r(q^i)
\end{equation}
where we have called the equation \(R(q)\). By design \(R(q)\) is smooth
enough to solve with Newton's method without any statemachine logic:
\begin{equation}
\frac{\partial R}{\partial q} \Delta q^{k+1} = - R(q^k)
\end{equation}
with the update
\begin{equation}
q^{k+1} = q^k + \alpha\Delta q^{k+1}
\end{equation}
where $\alpha$ is an under relaxation parameter.

This equation is deceptively simple: the function $R(q)$ has complex logic encoded in it; if-then
statements appear if the model used activations such as rectifiers.
However, the complex material behavior which traditionally requires
the vast majority of programming was implemented automatically by the
models; implementing equations XX, XX, and XX requires only a few lines
of human-written code.

\subsection{Dimensionalization}

Recall that the datasets are normalized to be zero-center and
unit-ranged to improve training. Thus, the outputs of the decoder are not in true
units.
The ranges are saved in a separate file which are loaded by the
simulation. The network is loaded, the $D(q)$ is extracted, and the following
transformations are applied to the outputs:
\begin{align}
T(q)      &= \left(T_{max}-T_{min}\right) D_T(q) +T_{mean} \\
p(q)      &= \exp\left(\left( p_{max}-p_{min}\right)
            D_p(q) +p_{mean} \right) \\
\rho(q) &= \left(\rho_{max}-\rho_{min}\right) D_{\rho}(q) +\rho_{mean}\\
h(q)      &= \left( h_{max}-h_{min}\right) D_h(q)  +h_{mean}
\end{align}
The exponent transformation is only applied to the cases where
$\log(p)$ was used to make the training set.
The range values are in the proper units, so that equation rows in the
simulation are in SI units.

\hypertarget{header-n3339}{%
\subsection{Initialization}\label{header-n3339}}

There is a caveat with representing the degrees of freedom of the system
on a learned space: the user does not know the meaning of $q$ as an
input. Further, the values of $q$ vary from network to network and
depend on the parameterization. The user needs to be able to specify the
"real" physical quantities.

Since the original manifold is 2D, the user only needs to specify two
initial values. In a multiphase solver such as TOUGH, the state needs
to be specified with the two primary variables plus the state tag (sans input file formatting): 
\begin{verbatim}
    9e5 20 Aqu
    0.4 20 AqG
\end{verbatim}
Note that the first primary variable means pressure in the first line
for a pure-liquid state,
but saturation in the second line for the liquid-gas equilibrium.
This makes sense to human intuition but
still leads to bugs and occasionally requires looking up values in
textbooks or the software manual. (Is it $S_{gas}$ or $S_{aqu}$ on the second
line?) A fully fledged solver needs to check the user specified
initial conditions for erroneous inputs, but it cannot detect ambiguities.

The extra process of determining the initial state means that more error checking is needed
in verifying user inputs, but this is necessary in any
multiphase/multicomponent reservoir simulator. LatentSim does not yet
detect and report errors, and only reports divergence. In our study, the
testing problem specifications are all correct, but not all network
realizations are acceptable. Failure to determine a \(q\) for one of the
test initial conditions is a common mode of detecting unacceptable
network realizations; this is marked as a testing error during training
which invalidates a model as a potential realization.

The current implementation LatentSim does not support inputting via
saturations on the equilibria, a commonly used primary variable, but
this problem could be solved with future engineering work for a
non-experimental simulator.

For LatentSim, the input system is not yet as robust as a
fully-fledged user-facing simulator, but the inputs are specified in
the $T p \rho h$ space instead of the $q$ space.
The user (or test problem) specifies any two of the four values in a
dictionary format,
\begin{verbatim}
    p=9e5 T=20
    rho=500 T=20
\end{verbatim}
Then, the simulator must determine $q(t=0)$ from these two
values. This sets up a system of two equations.

Note that the encoder only necessarily works when all four values are
correct. Thus, it is necessary to actually solve the nonlinear problem
on the decoder:
\begin{equation}
  D(q^0)_i = s^0_i \quad \text{for specified}\; i
\end{equation}
This equation has two rows indexing into the four state variables;
e.g. only the rows associated with $p$ and $T$ are solved if those are
specified, which is usually the case. The above equation is solved
with Newton's method. The two corresponding rows are cut out of
the tangent matrix to $D$,
\begin{equation}
K_{ij} = \frac{\partial D_i}{\partial q_j}  \quad \text{for
    specified}\; i, \text{for}\;j=1,2
\end{equation}

and the following system of equations is solved for an update to $q$,
\begin{equation}
K_{ij}\Delta q_j = s^0_i-D(q)_i
\end{equation}
The iteration on $q$ is repeated until convergence, with under-relaxation used again to update:
 \begin{equation}
q^0_{k+1} =q^0_{k} + \alpha \Delta q_{k}
\end{equation}
with the under-relaxation parameter usually set to $\alpha \approx
0.1$.

The initial condition problem is not guaranteed to converge, however,
and often does not.
For example, if the initial guess was in the liquid phase and the target point
in the EOS space is in a solid, the iteration will fail.
To address this issue, we perform multiple starts with an initial condition in each
phase and equilibrium, discard points that diverge, and then pick the
\(q\) that is closest to the two specified components. The multiple
start points are specified at EOS development time to form a set of
values that represent middle points for each phase,
\begin{equation}
\left\{s_{guess}\right\} = \left\{ \left(\begin{array}{c}
\bar{T}_{gas}\\
\bar{p}_{gas}\\
\bar{\rho}_{gas}\\
\bar{h}_{gas}
\end{array}\right),\left(\begin{array}{c}
\bar{T}_{liq}\\
\bar{p}_{liq}\\
\bar{\rho}_{liq}\\
\bar{h}_{liq}
\end{array}\right),\left(\begin{array}{c}
\bar{T}_{ice}\\
\bar{p}_{ice}\\
\bar{\rho}_{ice}\\
\bar{h}_{ice}
\end{array}\right)... \right\}
  \end{equation}
The multistart method could be specified without any prior knowledge
with random points, but this is a sufficiently robust method for
now. Each of the possible guess points is used to fill in the two
unspecified values as guesses and then passed into the encoder.
E.g. if $p$ and $T$ were specified, each possible $\bar{\rho}$ and
$\bar{h}$ combination from the above sets are filled in to provide
multiple guesses for $q$,
\begin{equation}
s^0 = \left(\begin{array}{c}
T^*(0)\\
p^*(0)\\
\bar{\rho}_{guess}\\
\bar{h}_{guess}
\end{array}\right);\quad q^0_{guess}=E\left(s^0\right)
 \end{equation}
 These sets of $q^0_{guess}$ are then iterated upon using the Newton's
 method procedure.

It is further possible that none of the initial guesses for this iteration
converge. There are two possibilities: the target state point is
incorrectly specified or out-of-range (e.g., negative density), or the
network is incapable of representing the desired material properties
(i.e., poorly trained). The first possibility shows an invalid
problem specification for the chosen EOS. The second possibility is
crucial to the automated training process. If a given network cannot
find an initial condition for every possible test problem, it fails
the tests and is marked as invalid. This is one way of automatically
verifying the training process.

\hypertarget{header-n3493}{%
\subsection{Methodology Summary}\label{header-n3493}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Make a database of \(T,p,\rho,h,X_1,...\)
  \begin{itemize}
  \item
    Piece together empirical fits for each phase from literature
  \item
    (\emph{Experimental data in the future})
  \end{itemize}
\item
  Normalize (and \(\log(p)\)) the database, shuffle it

  \begin{itemize}
  \item
    \(\log(p)\) distributes the low-\(p\) phases evenly w.r.t.
    high-\(p\) phases
  \end{itemize}
\item
  Train the autoencoder on batch-generated architectures. Note that
  phase labels are not used.
\item
  Load the models and generate physics code
\item
  Verify and grade architectures on tests

  \begin{itemize}
  \item
    Need more than autoencoder mean-squared-error
  \item
    Differentiability and numerical stability in simulation
  \item
    Evaluation speed (billions of times in a simulation!)
  \end{itemize}
\item
  Pick best one to package into production code
\end{enumerate}

\hypertarget{header-n3353}{%
\subsection{Pseudo-code}\label{header-n3353}}

\begin{verbatim}
load network from database
build DAE graph
input problem properties and initial value
solve for q0: D(q0) = s0
loop t=0 to t_max:
	solve for q[i]: R(q[i],q[i-1])=0
decode s[i] = D(q[i])
\end{verbatim}

\hypertarget{header-n3356}{%
\section{Implementation}\label{header-n3356}}

The training framework and simulations are implemented in TensorFlow 
using custom defined operations. The various visualizations are built
using Matplotlib, Plotly, and Visdom. The accompanying video was
generated using Paraview.

The codebase is released open source at
\url{https://github.com/afqueiruga/latentsim}. Helper methods and
custom optimizers that the
author reuses in various TensorFlow-based projects are factored out in
the dully-named \url{https://github.com/afqueiruga/afqstensorflowutils}.
The implementation of the empirical equations of state for water are
factored out at
\href{https://github.com/afqueiruga/equations_of_state}{https://github.com/afqueiruga/equations\_of\_state}.

\hypertarget{header-n3359}{%
\section{Simulation Evaluation}\label{header-n3359}}

Multiple hyperparameters are trained.

Each architecture trained on the autoencoder task is then

The phase space of water is broken up into three different extents to
change the complexity of the training process:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item Linear Equation of State: the range is $ p = 10^5+[-10^3, 10^3]
  Pa,\quad T = [ 19, 21 ] ^o C$. This reduces to the single phase
  Darcy's law problem where the material has constant compressibility
  and heat capacity. This is the most basic test of the training
  framework, and provides analytical solutions to test the simulation
  framework.
\item Liquid-Gas Regime: the range is  $ p = [100,5\times 10^5] Pa,
  \quad T = [274,594] K$. This restricts the domain to only one phase
  boundary. With this range, it is possible to find a linear
  combination of $T,p,\rho,h$ that can make a 1-1 mapping. (The usual
  $T-p$ combination does not, but $\rho-h$ would.)
\item Solid-Liquid-Gas-Supercritical Regimes: this covers the full
  extent of the parameterized equation of state of water,
    $ p = [6\times 10^{-6},3\times 10^8]Pa, \quad T = [150,1000] K
    $. This problem has no linear mapping to a unique latent space.
\end{enumerate}

\hypertarget{header-n3418}{%
\subsection{Validation and Testing}\label{header-n3418}}

The verification problems are listed in Table \ref{tab:problems}.
\begin{table}
  \caption{\label{tag:problems}Enumeration of the problems used to test and evaluate each fo the generated architectures}
\begin{tabular}{l|c|c|c}
  Name & start & path & end\\
\hline
\end{tabular}
\end{table}
The simulator was benchmarke against the long standing TOUGH+ multiphase
multicomponent

Single-grid block simulations were set up for most of the cases. TOUGH+
does not handle supercritical \ch{H2O}, so these test problems were not
used to benchmark. (The supercritical phase region is not practically
relevant for water, but it is very important for other fluids such as
\ch{CO2}.) (Note that, as a major motivation of this work, adding
supercritical states to the existing Fortran simulator would require
coding of one new logical state with two new state transitions with
primary variable switches; for the approach of this paper, the dataset
just needed to be extended.)

\hypertarget{header-n3421}{%
\section{Conclusions}\label{header-n3421}}

\begin{itemize}
\item
  Deep learning to replace and improve hand-baked equations and
  algorithms
\item
\item
  \textbf{Only training the material representation, not the balance
  laws}
\item
  Extend to more complicated materials
\item
  Put into a flow simulation
\item
  Close loop on testing with reinforcement learning
\end{itemize}

The ultimate goal for this work is to automate integration of data
analysis into simulation.

We hypothesize this end-to-end approach will allow the system to develop
\emph{better} simulations \emph{faster}.

We demonstrated a specially crafted architecture whose parameters can be
human interpretted and perform unsupervised phase/equilibria
classification.

We illustrated a programming model that is able to incorporate machine
learning models into existing well known computational methodologies.
This allows us to enforce physically known quantities, such as the
balance laws.

The TensorFlow system is not quite suited for this type of workflow -\/-
the resulting simulations are extremely slow due to repeated calls to
session.run and a very complicated graph structure for the polynomials.
Currently, the system is being reimplemented in Julia using Flux and
Zygote to seamlessly reincorporate into a fully-fledged new reservoir
simulator. The pendulum has been reimplemented in Julia at .

We are also demonstrating a new programming paradigm applied to
scientific computing. The entire system has another level of complexity
to it. An offline training phase searches for a new representation, i.e.
for new equations with new unknown degrees of freedom, that are good to
solve.

Each implemented equation of state (three options in this work spanning
different extents of \ch{H2O}) has a database of potential realizations.
The entries are model architectures, parameters, and training
environment configurations and logs for each equation-of-state dataset.
This database of models replaces a version controlled repository of user
written modules. Hundreds of trials for novel primary variable
configurations happen in a matter of minutes on a workstation in place
of pain-staking trial-and-error labor performed with human "intuition"
and experience. We did encode a little bit of human intuition into the
model architectures and tricks for initializing the networks.

We can also think about this work as offline preconditioning.

In TOUGH, for one instance, all of the calculations are performed in SI units because it is simply too much labor to normalize everything! Adverse effects to iterative algorithms and floating point error propogation occur, especially because numerical differentiation is used.
The learned primary variables are normalized to a unit range around zero, and the scaling matrix is explicitly in one step. The scaling matrix could be a way to systematically non-dimensionalize the entire problem.

The larger simulation code links into the trained database to load the
computer-written model into the computation graph for the simulation,
deriving all needed ingredients for the mainloop using automatic
differentiation.

The simulation inputs are the name of the equation of state, .e.g,
``water\_lg'' or ``water\_slgc'', and the name of the architecture
hyperparameters, e.g. ``Classifying\_pT\_1.0\_1,3,6,12,Sigmoid''.
In a mature code, the name of the architecture loaded would be hidden
from the end-user and would be a setting which the doman scientist and developer freezes
after checking the training process.

Techniquessuch as L1 regularization would be needed to search for low
cost models that also satisfy our desire for Occam's razor
simplifications. (The IAPWS relations used a prelimary form of this
type of simplification, but are not optimal in the author's opinion.)

A fully automated testing system can verify which realizations worked
(relate this to an independent review board comparing the results the
independently developed codes) and then select the fastest and most
robust one.

\section*{Acknowlegements}

This work was supported by Department of Energy, Office of Fossil
Energy contracts
1. Hydrates
2. Shales

The author would like to thank his colleagues Matt Reagan and George
Moridis for all their instruction on chemsitry. 
\bibliographystyle{plainnat}
\bibliography{zotero}

\end{document}